Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations

### 2025-03-21 index.Rmd
adding improves

```{r message=FALSE, warning=FALSE, echo=TRUE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # without S-R
dir4<-here("s4") # 
dir5<-here("s5") # 
dir6<-here("s6") # 
dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
dir8<-here("s8") # s1 platoons
dir9<-here("s9") # s1 w/ blocks
dirtest<-here("test")
Figs <- here("Figs")# S
```

@@

```{r}
# leo archivos para plotear y hacer tablas
start1 <- SS_readstarter(file = file.path(dir1,
                                          "starter.ss"),
                              verbose = FALSE)
# note the data and control file names can vary, so are determined from the 
# starter file.
dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
                        verbose = FALSE)
# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
# assumes the correct data structure
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
                                                "forecast.ss"),
                              verbose = FALSE)
# can also read in wtatage.ss for an empirical wt at age model using
# r4ss::SS_readwtatage()
```

@@

```{r}
parInit %>%
  kbl(booktabs = T,
      format = "html",
      position="ht!",
    caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=10,
                html_font = "arial")%>% 
  pack_rows(index = c("Mortalidad natural" = 1,
                        "Crecimiento"= 5,
                        "Relación longitud-peso" = 2,
                        "Ojiva de madurez"=2,
                        "Relación stock-recluta"=5,
                        "Capturabilidad"=2,
                        "Selectividad"=4))

@@

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1       |
|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s2    | "s1" without S-R relation                                                            |
|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
|    s6    | "s1" Ricker S-R relation estimated                                                   |
|    s7    | "s1" w/ set of parameters estimated in @EMM-204/32                                   |
|    s8    | "s1" test data weighting                                                             |
|    s9    | "s1" w/ selectivity blocks                                                             |

@@

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
# Lista de directorios para correro tordos juntos
directorios <- c("s01", 
                 "s1", 
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7",
                 "s8",
                 "s9")  # Agrega aquí todos los nombres de las carpetas que deseas procesar

# Bucle para ejecutar el código en cada directorio
for (dir in directorios) {
  r4ss::run(
    dir = dir,
    exe = "ss_osx",
    skipfinished = FALSE,
    show_in_console = TRUE
  )
}
```

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dirtest,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)
```

@@

```{r message=F, include=FALSE}
#s01
base.model01 <- SS_output(dir=dir01,
                        covar=T,
                        forecast=T)
#s1
base.model1 <- SS_output(dir=dir1,
                        covar=T,
                        forecast=T)
#s2
base.model2 <- SS_output(dir=dir2,
                        covar=T,
                        forecast=T)
#s3
base.model3 <- SS_output(dir=dir3,
                        covar=T,
                        forecast=T)
#s4
base.model4 <- SS_output(dir=dir4,
                        covar=T,
                        forecast=T)
#s5
base.model5 <- SS_output(dir=dir5,
                        covar=T,
                        forecast=T)
#s6
base.model6 <- SS_output(dir=dir6,
                        covar=T,
                        forecast=T)

#s7
base.model7 <- SS_output(dir=dir7,
                        covar=T,
                        forecast=T)

#s8
base.model8 <- SS_output(dir=dir8,
                        covar=T,
                        forecast=T)

#s9
base.model9 <- SS_output(dir=dir9,
                        covar=T,
                        forecast=T)
```

```{r message=F, include=FALSE}
#stest
base.modeltest <- SS_output(dir=dirtest,
                        covar=T,
                        forecast=T)
```

Just some time series plots

```{r}
SSplotTimeseries(base.modeltest,
                 subplot = 7)
SSplotTimeseries(base.modeltest,
                 subplot = 9)
```
and fits

```{r}
SSplotComps(base.modeltest,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```

```{r eval= FALSE, message=F, include=FALSE}
# html
SS_plots(base.modeltest,
         uncertainty=T,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75,
         minbthresh=0.20, 
         forecast=T)

```
# RESULTS

@@

```{r eval=FALSE, message=F, include=FALSE}
SS_plots(base.model7, 
         uncertainty = TRUE,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.2, 
         forecast=T)
```

@@

```{r }
SSplotData(base.modeltest, subplot = 1)

SSplotData(base.model1, 
           subplot = 1,
           pheight = 15)
```

@@

```{r fig.height=6}
SSplotSelex(base.model01,
            subplots = 1)
SSplotSelex(base.model1,
            subplots = 1)
SSplotSelex(base.model2,
            subplots = 1)
SSplotSelex(base.model3,
            subplots = 1)
SSplotSelex(base.model4,
            subplots = 1)
SSplotSelex(base.model5,
            subplots = 1)
SSplotSelex(base.model6,
            subplots = 1)
SSplotSelex(base.modeltest,
            subplots = 1)
```

@@

```{r}
SSplotRunstest(base.modeltest,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r eval=FALSE, fig.width=4, fig.height=5}
SSplotRunstest(base.model01,
               subplots = "len",
               add=T)
SSplotRunstest(base.model1,
               subplots = "len",
               add=T)
SSplotRunstest(base.model2,
               subplots = "len",
               add=T)
SSplotRunstest(base.model3,
               subplots = "len",
               add=T)
SSplotRunstest(base.model4,
               subplots = "len",
               add=T)
SSplotRunstest(base.model5,
               subplots = "len",
               add=T)
SSplotRunstest(base.model6,
               subplots = "len",
               add=T)
SSplotRunstest(base.modeltest,
               subplots = "len",
               add=T)
```

```{r}
SSplotRunstest(base.modeltest,
               subplots = "cpue",
               add=T)
```

@@

```{r eval=FALSE}
SSplotJABBAres(base.model01,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "len",
               add=T)
SSplotJABBAres(base.modeltest,
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model01,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.modeltest,
               subplots = "cpue",
               add=T)
```

## Relationship Stock-Recruit

```{r eval=FALSE}
l <- base.model1$recruit$SpawnBio
m <- base.model1$recruit$pred_recr
ggplot()+
  aes(l,m)+
  geom_point()+
  geom_smooth(method = "loess", 
              se =F)+
  theme_few()+
  ylim(0, max(base.model1$recruit$pred_recr))
```

@@

```{r eval=FALSE}
#one by one
retro(
    dir = dirtest,
    oldsubdir = "",
    newsubdir = "Retrospective",
    years = 0:-4,
    exe = "ss_osx",
    extras = "-nox",
    skipfinished = FALSE)
```

@@

```{r eval=FALSE, message=FALSE, warning=FALSE}
#stest
retroModelstest <- SSgetoutput(dirvec=file.path(dirtest,
                                            "Retrospective",
                                            paste("retro",0:-4,
                                                  sep="")))

retroSummarytest <- SSsummarize(retroModelstest)

@@

retrotest <- SSplotRetro(retroSummarytest,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

```{r eval=FALSE}
SSplotRetro(retroSummarytest,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
```

@@

mvln = SSdeltaMVLN(base.model7,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

>>>>

```{r message=FALSE, warning=FALSE, echo=TRUE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # without S-R
dir4<-here("s4") # 
dir5<-here("s5") # 
dir6<-here("s6") # 
dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
dir8<-here("s8") # s1 platoons
dir9<-here("s9") # s1 w/ blocks
dir1.1<-here("s1.1") # con todo
dir1.2 <- here("s1.2") # sin predador ni ambiental
dir1.3 <- here("s1.3") # borrado los registros de prreddador )igual q el S2
Figs <- here("Figs")# S
```

@@

```{r}
# leo archivos para plotear y hacer tablas
start1 <- SS_readstarter(file = file.path(dir1.1,
                                          "starter.ss"),
                              verbose = FALSE)
# note the data and control file names can vary, so are determined from the 
# starter file.
dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
                        verbose = FALSE)
# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
# assumes the correct data structure
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
                                                "forecast.ss"),
                              verbose = FALSE)
# can also read in wtatage.ss for an empirical wt at age model using
# r4ss::SS_readwtatage()
```

@@

```{r}
parInit %>%
  mutate(across(where(is.numeric), ~ round(., 3)))  %>% 
  kbl(booktabs = TRUE,
      format = "html",
      position = "ht!",
      caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", full_width = TRUE) %>%  # Expande el ancho
  kable_styling(latex_options = c("striped", "condensed"),
                full_width = TRUE,  # Expande la tabla al ancho completo
                font_size = 12,
                html_font = "Times New Roman") %>%  # Aplica la fuente Times New Roman
  pack_rows(index = c("Mortalidad natural" = 1,
                      "Crecimiento" = 5,
                      "Relación longitud-peso" = 2,
                      "Ojiva de madurez" = 2,
                      "Relación stock-recluta" = 5,
                      "Capturabilidad" = 2,
                      "Selectividad" = 4))

@@

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1       |
|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s2    | "s1" without S-R relation                                                            |
|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
|    s6    | "s1" Ricker S-R relation estimated                                                   |
|    s7    | "s1" w/ set of parameters estimated in @EMM-204/32                                   |
|    s8    | "s1" test data weighting                                                             |
|    s9    | "s1" w/ selectivity blocks                                                           |
|    s1.1  | Spatial data without environmental and predator components                          |
|    s1.2  | "s1.1" with environmental and predator components                                    |
|    s1.3  | "s1.2" w/ selectivity blocks                                                         |

@@

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
directorios <- c("s1.1",
                 "s1.2",
                 "s1.3")

for (dir in directorios) {
  r4ss::run(
    dir = dir,
    exe = "ss_osx",
    skipfinished = FALSE,
    show_in_console = TRUE
  )
}
```

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir1.1,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

base.model1.1 <- SS_output(dir=dir1.1,
                        covar=T,
                        forecast=T)
```

@@

```{r message=F, include=FALSE}

#s1.1
base.model1.1 <- SS_output(dir=dir1.1,
                        covar=T,
                        forecast=T)
#s1.2
base.model1.2 <- SS_output(dir=dir1.2,
                        covar=T,
                        forecast=T)
#s1.3
base.model1.3 <- SS_output(dir=dir1.3,
                        covar=T,
                        forecast=T)
```

```{r}
SSplotTimeseries(base.model1.1,
                 subplot = 1)
SSplotTimeseries(base.model1.1,
                 subplot = 9)
SSplotTimeseries(base.model1.1,
                 subplot = 7)
```
and fits

```{r}
SSplotBiology(base.model1.1,
            subplots = 1:32)
```

```{r}
SSplotComps(base.model1.1,
            subplots = 1)

# RESULTS
```{r eval=FALSE, message=F, include=FALSE}
# Definir la ruta de la carpeta
folder_path <- "s1.1/plots"
# Verificar si la carpeta existe y eliminarla
if (dir.exists(folder_path)) {
  unlink(folder_path, recursive = TRUE)
  message("La carpeta 'plots' ha sido eliminada.")
} else {
  message("La carpeta 'plots' no existe.")
}

SS_plots(base.model1.1, 
         uncertainty = TRUE,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.2, 
         forecast=T)
```

@@

@@

```{r }
SSplotData(base.model1.1, 
           subplot = 1,
           pheight = 15)
```

@@

```{r fig.height=6}
SSplotSelex(base.model01,
            subplots = 1)
SSplotSelex(base.model1,
            subplots = 1)
SSplotSelex(base.model2,
            subplots = 1)
SSplotSelex(base.model3,
            subplots = 1)
SSplotSelex(base.model4,
            subplots = 1)
SSplotSelex(base.model5,
            subplots = 1)
SSplotSelex(base.model6,
            subplots = 1)
SSplotSelex(base.model1.1,
            subplots = 1)
```

@@

```{r}
SSplotRunstest(base.model1.1,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r eval=FALSE}
SSplotRunstest(base.model01,
               subplots = "len",
               add=T)
SSplotRunstest(base.model1,
               subplots = "len",
               add=T)
SSplotRunstest(base.model2,
               subplots = "len",
               add=T)
SSplotRunstest(base.model3,
               subplots = "len",
               add=T)
SSplotRunstest(base.model4,
               subplots = "len",
               add=T)
SSplotRunstest(base.model5,
               subplots = "len",
               add=T)
SSplotRunstest(base.model6,
               subplots = "len",
               add=T)
SSplotRunstest(base.modeltest,
               subplots = "len",
               add=T)
```

```{r}
SSplotRunstest(base.model1.1,
               subplots = "cpue",
               add=T)
```

@@

```{r eval=FALSE}
SSplotJABBAres(base.model01,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1.1,
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model01,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1.1,
               subplots = "cpue",
               add=T)
```
## Relationship Stock-Recruit

```{r eval=FALSE}
l <- base.modeltest$recruit$SpawnBio
m <- base.modeltest$recruit$pred_recr
ggplot() +
  aes(l, m) +
  geom_point() +
  geom_smooth(method = "loess",
              se = TRUE, 
              color = "blue") +
  theme_few() +
  ylim(0, max(base.modeltest$recruit$pred_recr))

@@

```{r eval=FALSE}
#one by one
retro(
    dir = dir1.1,
    oldsubdir = "",
    newsubdir = "Retrospective",
    years = 0:-4,
    exe = "ss_osx",
    extras = "-nox",
    skipfinished = FALSE)
```

@@

```{r eval=FALSE, message=FALSE, warning=FALSE}
#stest
retroModels1.1 <- SSgetoutput(dirvec=file.path(dir1.1,
                                            "Retrospective",
                                            paste("retro",0:-4,
                                                  sep="")))

retroSummary1.1 <- SSsummarize(retroModels1.1)

@@

retro1.1 <- SSplotRetro(retroSummary1.1,
            add=T,
            forecast = F,
            legend = T,
            verbose=F,
            subplots = c("SSB"))
```

@@

mvln = SSdeltaMVLN(base.model1.1,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)
